validateHOT - Validate your Holdout Task

### 2023-10-26 paper/paper.Rmd
updated paper - 231025

---
title: 'validateHOT - Validate your Holdout Task'
tags:
  - R
  - MaxDiff
  - Conjoint Analysis
  - Market Simulations
  - Predictive Validity
authors:
  - name: Joshua Schramm
    orcid: 0000-0001-5602-4632
    corresponding: True
    affiliation: 1
  - name: Marcel Lichters
    orcid: 0000-0002-3710-2292
    corresponding: FALSE
    affiliation: 1, 2    
affiliations:
 - name: Chemnitz University of Technology, Germany
   index: 1
 - name: Otto von Guericke University of Magdeburg, Germany
   index: 2
citation_author: Schramm & Lichters.
date: 22 October 2023
year: 2023
bibliography: paper.bib
link-citations: true
output: rticles::joss_article
journal: JOSS
---

\setlength{\headheight}{63.55022pt}
\addtolength{\topmargin}{-0.95425pt}
\newcommand{\colcod}[1]{\texttt{\color{purple}#1}}

```{r include=FALSE}
library(validateHOT)
library(tidyverse)
```

@@

validateHOT is a package that provides functions to both validate a validation/holdout task and run market simulations for results obtained in a (adaptive) choice-based conjoint analysis (hereafter ACBC and CBC, respectively) and maximum difference scaling (hereafter MaxDiff) using [Sawtooth Software](https://sawtoothsoftware.com/). The ultimate goal of preference measurement techniques such as (A)CBC or MaxDiff is to predict future behavior [@green1990]. Therefore, it is essential for both academics and practitioners to ensure that the collected data is valid and can also predict outside tasks. The easiest way to test this is to include so-called holdout or validation task [@Orme2015]. Despite the important role of validation tasks, most of the conjoint studies done in practice do not include them [@yang2018], which is unsatisfactory, given the fact that the model is used to estimate market shares which is the basis for relevant marketing decisions.

@@

validateHOT is a practical tool for Sawtooth Software users in industry as well as academia. It provides an open source solution for a) validating a validation/holdout task and therefore ensuring that the model has predictive validity; b) running market simulations. Other packages, for example, Metrics [@Metrics] provide functions to run validation metrics such as *mean absolute error*, *root mean squared error*, or the five metrics of the confusion matrix (see \autoref{tab:table1}). However, to put the Sawtooth export into the right format, the user needs some data wrangling which can be a barrier. Other packages mainly focus on the analysis of conjoint analysis (e.g., ChoiceModelR [@ChoiceModelR], choicetools [@choicetools], logitR [@logitr], bayesm [@bayesm] etc.). To the best of our knowledge, a package that converts raw utility scores into validation metrics or running a variety of marketing simulations is still missing. The goal of validateHOT is still to fill this research gap. validateHOT creates market shares based on *share of preferences* or *first choice rule*, which can be helpful to report results for research articles according to open science standards. In addition, it also provides the function \colcod{turf}, which was offered formerly by the turfR package, which is no longer available on the Comprehensive R Archive Network (see [CRAN](https://cran.r-project.org/web/packages/turfR/index.html)). So, currently practitioners and academics mainly have to stick to paid solutions, for example, Sawtooth Software.

@@

After running the Hierarchical Bayes estimation in Sawtooth Software, the raw utility scores should be exported. Both the raw utilities and the actual choice in the validation/holdout task needs be read into *R* (please see the vignette for more advice). In the following, we provide the workflow for a MaxDiff study (the vignette also provides detailed examples for a CBC as well as an ACBC).

We create the validation/holdout task, which has a total of 7 items (\texttt{\color{purple}prod}) plus the no-buy alternative (\texttt{\color{purple}none}). Moreover, we also define the method, the actual choice in the validation/holdout task as well as the column indexes of the alternatives in the validation/holdout task. Finally, we define variables that we would like to keep in the data frame (\colcod{varskeep}). \texttt{\color{purple}createHOT} returns a data frame which we need for subsequent analyses.

@@

Next, validateHOT also provides the five metrics for the confusion matrix. The underlying logic hereby is that the user has to provide a no-buy alternative (\texttt{\color{purple}none}). validateHOT calculates, for example, how often a buy or no-buy was correctly predicted, therefore, it is testing whether the model correctly predicts general demand (exemplary showed by applying the \texttt{\color{purple}accuracy} function and results split by \texttt{\color{purple}Group}).

@@

For the following example, let us assume that you conducted an anchored MaxDiff analysis with 10 items (\colcod{opts}) and you want to find the best assortment with a size of 3 (\colcod{size = 3}). As a threshold (\colcod{none}) you use the anchor (no-buy alternative). Finally, you apply the threshold approach, that is that each alternative with a utility above the threshold (in our case \colcod{none}) is marked as purchase option. We reduce the output to the Top 5 combos based on the reach (a \colcod{1} indicates present in this assortment while \colcod{0} means not present).

```{r}
turf(
  data = MaxDiff,
  opts = c(Option_01:Option_10),
  none = none,
  size = 3,
  approach = "thres"
) %>% 
  head(., n = 5) %>% 
  mutate_if(is.numeric, round, 2) %>% 
  t() %>% 
  as.data.frame() %>% 
  slice(-1) %>% 
  rename_all(., ~ paste0("Combo ", c(1:5)))
```

>>>>

---
title: 'validateHOT - Validate your Holdout Task'
tags:
  - R
  - MaxDiff
  - Conjoint Analysis
  - Market Simulations
  - Predictive Validity
authors:
  - name: Joshua Schramm
    orcid: 0000-0001-5602-4632
    corresponding: True
    affiliation: 1
  - name: Marcel Lichters
    orcid: 0000-0002-3710-2292
    corresponding: FALSE
    affiliation: 1, 2    
affiliations:
 - name: Chemnitz University of Technology, Germany
   index: 1
 - name: Otto von Guericke University of Magdeburg, Germany
   index: 2
citation_author: Schramm & Lichters
date: 22 October 2023
year: 2023
bibliography: paper.bib
link-citations: true
output: rticles::joss_article
journal: JOSS
---

```{=tex}
\setlength{\headheight}{63.55022pt}
\addtolength{\topmargin}{-0.95425pt}
\newcommand{\colcod}[1]{\texttt{\color{purple}#1}}
```
```{r include=FALSE}
library(validateHOT)
library(tidyverse)
```

@@

validateHOT is a package that provides functions to both validate a validation/holdout task and run market simulations for results obtained in a (adaptive) choice-based conjoint analysis (hereafter ACBC and CBC, respectively) and maximum difference scaling (hereafter MaxDiff) using [Sawtooth Software](https://sawtoothsoftware.com/).

Preference measurement techniques', such as (A)CBC or MaxDiff, ultimate goal is to predict future behavior [@green1990]. Hence, it is essential for both academics and practitioners to ensure that the collected data is valid and can also predict outside tasks. The easiest way to test this is to include so-called holdout or validation task [@Orme2015]. Despite the important role of validation tasks, practictioners often do not include them [@yang2018], which is unsatisfactory, given the fact that the model is used to estimate market shares which is the basis for relevant marketing decisions.

@@

validateHOT is a practical tool for Sawtooth Software users in industry as well as academia. It provides an open source solution for a) validating a validation/holdout task and therefore ensuring that the model has predictive validity; b) running market simulations. Other packages, for example, Metrics [@Metrics] provide functions to run validation metrics such as *mean absolute error*, *root mean squared error*, or the five metrics of the confusion matrix (see \autoref{tab:table1}). However, to put the Sawtooth export into the right format, the user needs some data wrangling which can be a barrier. Other packages mainly focus on the analysis of conjoint analysis (e.g., ChoiceModelR [@ChoiceModelR], choicetools [@choicetools], logitR [@logitr], bayesm [@bayesm] etc.). To the best of our knowledge, a package that converts raw utility scores into validation metrics or running a variety of marketing simulations is still missing. The goal of validateHOT is still to fill this research gap. validateHOT creates market shares based on *share of preferences* or *first choice rule*, which can be helpful to report results for research articles according to open science standards. In addition, it also provides the function \texttt{\color{purple}turf}, which was offered formerly by the turfR package, which is no longer available on the Comprehensive R Archive Network (see [CRAN](https://cran.r-project.org/web/packages/turfR/index.html)). So, currently practitioners and academics mainly have to stick to paid solutions, for example, Sawtooth Software.

@@

In the following, we provide the workflow for a MaxDiff study (the vignette also provides detailed examples for a CBC as well as an ACBC).

After running the Hierarchical Bayes estimation in Sawtooth Software, the **raw** utility scores have to be exported and afterwards read into *R* including the actual choice in the validation/holdout task.

To define the validation/holdout task, which has a total of 7 items (\texttt{\color{purple}prod}) plus the no-buy alternative (\texttt{\color{purple}none}), we use the \texttt{\color{purple}createHOT} function. Here, the user can define the attributes as well as the method (in this case \texttt{\color{purple}MaxDiff}).

@@

validateHOT also provides the five metrics for the confusion matrix. The underlying logic hereby is that the user has to provide a no-buy alternative (\texttt{\color{purple}none}). validateHOT calculates, for example, how often a buy or no-buy was correctly predicted, therefore, it is testing whether the model correctly predicts general demand (exemplary showed by applying the \texttt{\color{purple}accuracy} function and results split by \texttt{\color{purple}Group}).

@@

For the following example, let us assume that you conducted an anchored MaxDiff analysis with 10 items (\texttt{\color{purple}opts}) and you want to find the best assortment with a size of 3 (\texttt{\color{purple}size = 3}). As a threshold (\texttt{\color{purple}none}) you use the anchor (no-buy alternative).

```{r}
turf(
  data = MaxDiff, # define data
  opts = c(Option_01:Option_10), # define items
  none = none, # define threshold variable
  size = 3, # define size of assortment
  approach = "thres" # define approach
) %>%
  head(., n = 5) %>%
  mutate_if(is.numeric, round, 2) %>%
  t() %>%
  as.data.frame() %>%
  slice(-1) %>%
  rename_all(., ~ paste0("Combo ", c(1:5)))
```
